---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sigshared

<!-- badges: start -->

[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN
status](https://www.r-pkg.org/badges/version/sigshared)](https://CRAN.R-project.org/package=sigshared)
[![R-CMD-check](https://github.com/selkamand/sigshared/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/selkamand/sigshared/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

**sigshared** contains general utilities used in many other R packages
in the [sigverse](https://github.com/selkamand/sigverse)

Unless you're developing / maintaining a sigverse-compatible package,
there is probably very little reason to ever download this package
explicitly. We recommend just installing the full sigverse as described
[here](https://github.com/selkamand/sigverse)

## Installation

You can install the development version of sigshared from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("selkamand/sigshared")
```

## Sigverse Data Types

+------------------------------+------------------------------+
| Data Structure               | Requirements                 |
+==============================+==============================+
| **Signature**                | The profile of a mutational  |
|                              | signature                    |
|                              |                              |
|                              | data.frames with 3 columns   |
|                              |                              |
|                              | 1.  **channel**              |
|                              | 2.  **type**                 |
|                              | 3.  **fraction**             |
+------------------------------+------------------------------+
| **Signature Collections**    | Lists of signature           |
|                              | data.frames, where name of   |
|                              | list entry is the name of    |
|                              | the signature                |
+------------------------------+------------------------------+
| **Signature Annotations**    | Signature level annotations. |
|                              | data.frames with 4 required  |
|                              | columns:                     |
|                              |                              |
|                              | 1.  **signature**            |
|                              | 2.  **aetiology**            |
|                              | 3.  **class**                |
|                              | 4.  **subclass**             |
|                              |                              |
|                              | class and subclass of        |
|                              | aetiology do not need to     |
|                              | conform to any specific      |
|                              | ontology. However, we        |
|                              | include the data-dictionary  |
|                              | used by sigstash collections |
|                              | below (see [Signature        |
|                              | Aetiology Classes])          |
+------------------------------+------------------------------+
| **Catalogue**                | The mutational profile of a  |
|                              | sample, described by         |
|                              | tallying mutations belonging |
|                              | to each mutational channel.  |
|                              | Catalogues are not always    |
|                              | observational. They can also |
|                              | be simulated from signature  |
|                              | models. data.frames with 4   |
|                              | required columns:            |
|                              |                              |
|                              | 1.  **channel**              |
|                              |                              |
|                              | 2.  **type**                 |
|                              |                              |
|                              | 3.  **fraction**             |
|                              |                              |
|                              | 4.  **count**                |
+------------------------------+------------------------------+
| **Catalogue Collections**    | Lists of catalogue           |
|                              | data.frames (1 per sample)   |
|                              | where name represents a      |
|                              | sample identifier.           |
+------------------------------+------------------------------+
| **Cohort Signature Analysis  | data.frame with 4 columns:   |
| Results**                    |                              |
|                              | 1.  **sample**               |
|                              |                              |
|                              | 2.  **signature**            |
|                              |                              |
|                              | 3                            |
|                              | .  **contribution_absolute** |
|                              |                              |
|                              | 4.  **contribution**         |
|                              |                              |
|                              | 5.  **p_value** (see         |
|                              |     \`?sigstats::sig_com p   |
|                              |                              |
|                              | u                            |
|                              | te_experimental_p_value()\`) |
+------------------------------+------------------------------+
| **Bootstraps**               | data.frame with 1 row per    |
|                              | signature per bootstrap      |
|                              |                              |
|                              | 1.  **bootstrap**            |
|                              | 2.  **signature**            |
|                              | 3                            |
|                              | .  **contribution_absolute** |
|                              | 4.  **contribution**         |
|                              |     (percentage)             |
+------------------------------+------------------------------+
| **Model Specification**      | Named numeric vector where   |
|                              | names represent signatures   |
|                              | and values represent their   |
|                              | proportional contribution to |
|                              | the model.                   |
+------------------------------+------------------------------+
| **Cohort Metadata**          | Data frames describing       |
|                              | sample-level metadata with   |
|                              | required columns:            |
|                              |                              |
|                              | 1.  **sample**               |
|                              | 2.  **disease**              |
|                              |                              |
|                              | Can include additional       |
|                              | columns with other metadata. |
+------------------------------+------------------------------+
| **UMAP**                     | Data frames representing     |
|                              | UMAP coordinates with at     |
|                              | least 3 columns:             |
|                              |                              |
|                              | 1.  **sample**               |
|                              | 2.  **dim1**                 |
|                              | 3.  **dim2**                 |
+------------------------------+------------------------------+
| **Similarity Against         | data.frame that describes    |
| Cohort**                     | how similar a sample         |
|                              | catalogue is to others in    |
|                              | the cohort. Contains 2       |
|                              | columns:                     |
|                              |                              |
|                              | 1.  **sample**               |
|                              | 2.  **cosine_similarity**    |
+------------------------------+------------------------------+

### Assertions

You can assert an object belongs to each of data types

```{r}
library(sigshared)

# Generate Example Datatypes

# Signatures
signature = example_signature()
signature_collection = example_signature_collection()
signature_annotations = example_annotations()

# Catalogues
catalogue = example_catalogue()
catalogue_collection = example_catalogue_collection()

# Cohort Analysis Results
cohort_analysis = example_cohort_analysis()

# Model
model = example_model()

# Assert Signatures
assert_signature(signature)
assert_signature_collection(signature_collection)
assert_signature_annotations(signature_annotations)

# Assert catalogues
assert_catalogue(catalogue)
assert_catalogue_collection(catalogue_collection)

# Assert Analyses
assert_cohort_analysis(cohort_analysis)

# Assert Model
assert_model(model, signature_collection,arg_name = "bob")
```

### Signature Aetiology Classes {data-link="Signature Aetiology Classes"}

```{r warning=FALSE, results='asis'}
library(knitr)
kable(sig_aetiology_classes())
```

### For Developers

#### Argument Naming Conventions

- **`catalogue`**  
  A single mutational catalogue for a sample. A `data.frame` with columns: `channel`, `type`, `fraction`, `count`. Catalogues may be empirical (observed) or simulated.  
  _See_: `sigshared::example_catalogue()`

- **`signature`**  
  A mutational signature profile. A `data.frame` with columns: `type`, `channel`, `fraction`.  
  _See_: `sigshared::example_signature()`

- **`signatures`**  
  A collection of signatures. A named list of `signature` data.frames.  
  _See_: `sigshared::example_signature_collection()`

- **`catalogues`**  
  A collection of catalogues. A named list of `catalogue` data.frames.  
  _See_: `sigshared::example_catalogue_collection()`

- **`model`**  
  A named numeric vector describing a signature mixture. Names are signature IDs, values are their proportional contributions (e.g., `c(SBS1 = 0.6, SBS5 = 0.4)`).  
  _See_: `sigshared::example_model()`

- **`cohort`**  
  A data.frame describing signature contributions per sample. Columns: `sample`, `signature`, `contribution_absolute`, `contribution`.  
  _See_: `sigshared::example_cohort()`

- **`cohort_metadata`**  
  Sample-level metadata as a data.frame. Must include columns: `sample`, `disease`. Can include others.  
  _See_: `sigshared::example_metadata()`

- **`similarity_against_cohort`**  
  A data.frame summarizing pairwise similarity between a sample and all others in the cohort. Columns: `sample`, `cosine_similarity`.  
  _See_: `sigshared::example_similarity_against_cohort()`

- **`umap`**  
  A 2D UMAP projection of catalogue similarities. A data.frame with columns: `sample`, `dim1`, `dim2`.  
  _See_: `sigshared::example_umap()`

## Example Data

For each data structure used by the sigverse we include a toy example.

```{r}
example_annotations()
example_bootstraps()

example_catalogue()
example_catalogue_collection()

example_signature()
example_signature_collection()

example_model()

example_cohort_analysis()
example_similarity_against_cohort()

example_cohort_metadata()

example_umap()
```

We also include examples from real SBS mutational signature analysis of
the colo829 melanoma cell line

```{r paged.print=TRUE}
head(example_catalogue_colo829())
head(example_bootstraps_colo829())
```

## Reformatting Signature/Catalogue Collections

```{r}
# List of signatures -> matrix (rows = channels; columns = signatures; values = fractions)
sig_collection_reformat_list_to_matrix(example_signature_collection())

# Matrix -> List of signatures
sig_collection_reformat_matrix_to_list(example_signature_collection_matrix(),
                                       values = "fraction")

# List of signatures -> tidy data.frame
sig_collection_reformat_list_to_tidy(example_signature_collection())

# Tidy data.frame -> list of signatures
sig_collection_reformat_tidy_to_list(example_signature_collection_tidy())



# All the above methods work with catalogues
sig_collection_reformat_list_to_matrix(example_catalogue_collection(), values = "count")

sig_collection_reformat_matrix_to_list(example_catalogue_collection_matrix(), 
                                       values = "count")

```

## Other Utility Functions

sigverse dependencies are kept minimal. To mitigate the cost to
readability, sigshared includes some baseR implementations of common
data.frame manipulation functions and other utilities.

```{r}
# Setup data
mtcars <- mtcars[1:5,1:5]

# Rename a dataframe
brename(mtcars, c(miles_per_gallon = "mpg"))

# Select a subset of columns
bselect(mtcars, c("mpg"))

# Evaluate code with a specific random seed
with_seed(seed = 123, { runif(1) })

# Compute fraction from count vector
compute_fraction(c(1, 100, 10, 40))
```

## S3 classes

Most types of data used by sigverse are so simple-stuctured we can just
expect data.frames/lists and use custom assertions to ensure it matches
expectation. There are a couple of exceptions, however, where we provide
S3 classes abstract away complexity in the datastores. For example we
use 2 different s3 objects to store signature analysis results.

1.  **signature_analysis_results**\
    The numeric results of signature analysis (no-visualisation).
    Designed to allow a creation of all sigstory visualisations from
    EXCLUSIVELY data in the object.

2.  **sigstory_visualisations**\
    Contains mostly visualisations only + metrics we display in sigstory
    reports. This helps us keep sigstory very light (logic is hard to
    debug in knitted quarto templates). Additionally, this object type
    allows us to easily write all visualisations to disk so they can be
    pulled into non-sigstory reporting tools.

Unless you're developing a sigstory compatible R package you probably
won't ever need to think about either of these object types.
