---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sigshared

<!-- badges: start -->

[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN
status](https://www.r-pkg.org/badges/version/sigshared)](https://CRAN.R-project.org/package=sigshared)
[![R-CMD-check](https://github.com/selkamand/sigshared/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/selkamand/sigshared/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

**sigshared** contains general utilities used in many other R packages
in the [sigverse](https://github.com/selkamand/sigverse)

Unless you're developing / maintaining a sigverse package, there is
probably very little reason to ever download this package explicitly. We
reccomend just installing the full sigverse as described
[here](https://github.com/selkamand/sigverse)

## Installation

You can install the development version of sigshared from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("selkamand/sigshared")
```

## Sigverse Data Types

+--------------------------------+--------------------------------+
| Data Structure                 | Requirements                   |
+================================+================================+
| **Signature**                  | The profile of a mutational    |
|                                | signature                      |
|                                |                                |
|                                | data.frames with 3 columns     |
|                                |                                |
|                                | 1.  **channel**                |
|                                | 2.  **type**                   |
|                                | 3.  **fraction**               |
+--------------------------------+--------------------------------+
| **Signature Collections**      | Lists of signature             |
|                                | data.frames, where name of     |
|                                | list entry is the name of the  |
|                                | signature                      |
+--------------------------------+--------------------------------+
| **Signature Annotations**      | Signature level annotations.   |
|                                | data.frames with 4 required    |
|                                | columns:                       |
|                                |                                |
|                                | 1.  **signature**              |
|                                | 2.  **aetiology**              |
|                                | 3.  **class**                  |
|                                | 4.  **subclass**               |
|                                |                                |
|                                | class and subclass of          |
|                                | aetiology do not need to       |
|                                | conform to any specific        |
|                                | ontology. However, we include  |
|                                | the data-dictionary used by    |
|                                | sigstash collections below     |
|                                | (see [Signature Aetiology      |
|                                | Classes])                      |
+--------------------------------+--------------------------------+
| **Catalogue**                  | The mutational profile of a    |
|                                | sample, described by tallying  |
|                                | mutations belonging to each    |
|                                | mutational channel. Catalogues |
|                                | are not always observational.  |
|                                | They can also be simulated     |
|                                | from signature models.         |
|                                | data.frames with 4 required    |
|                                | columns:                       |
|                                |                                |
|                                | 1.  **channel**                |
|                                |                                |
|                                | 2.  **type**                   |
|                                |                                |
|                                | 3.  **fraction**               |
|                                |                                |
|                                | 4.  **count**                  |
+--------------------------------+--------------------------------+
| **Catalogue Collections**      | Lists of catalogue data.frames |
|                                | (1 per sample) where name      |
|                                | represents a sample            |
|                                | identifier.                    |
+--------------------------------+--------------------------------+
| **Cohort Signature Analysis    | data.frame with 4 columns:     |
| Results**                      |                                |
|                                | 1.  **sample**                 |
|                                |                                |
|                                | 2.  **signature**              |
|                                |                                |
|                                | 3.  **contribution_absolute**  |
|                                |                                |
|                                | 4.  **contribution**           |
|                                |                                |
|                                | 5.  **p_value** (see           |
|                                |     \`?sigstats::sig_com p     |
|                                |                                |
|                                | ute_experimental_p_value()\`)  |
+--------------------------------+--------------------------------+
| **Bootstraps**                 | data.frame with 1 row per      |
|                                | signature per bootstrap        |
|                                |                                |
|                                | 1.  **bootstrap**              |
|                                | 2.  **signature**              |
|                                | 3.  **contribution_absolute**  |
|                                | 4.  **contribution**           |
|                                |     (percentage)               |
+--------------------------------+--------------------------------+
| **Model Specification**        | Named numeric vector where     |
|                                | names represent signatures and |
|                                | values represent their         |
|                                | proportional contribution to   |
|                                | the model.                     |
+--------------------------------+--------------------------------+

### Assertions

You can assert an object belongs to each of data types

```{r}
library(sigshared)

# Generate Example Datatypes

# Signatures
signature = example_signature()
signature_collection = example_signature_collection()
signature_annotations = example_annotations()

# Catalogues
catalogue = example_catalogue()
catalogue_collection = example_catalogue_collection()

# Cohort Analysis Results
cohort_analysis = example_cohort_analysis()

# Model
model = example_model()

# Assert Signatures
assert_signature(signature)
assert_signature_collection(signature_collection)
assert_signature_annotations(signature_annotations)

# Assert catalogues
assert_catalogue(catalogue)
assert_catalogue_collection(catalogue_collection)

# Assert Analyses
assert_cohort_analysis(cohort_analysis)

# Assert Model
assert_model(model, signature_collection,arg_name = "bob")
```

### Signature Aetiology Classes {data-link="Signature Aetiology Classes"}

```{r warning=FALSE, results='asis'}
library(knitr)
kable(sig_aetiology_classes())
```

### For Developers

#### Argument Naming

*Catalogue*

```         
catalogue: The mutational profile of a sample, described by tallying mutations belonging to each mutational channel. Catalogues are not always observational. They can also be simulated from signature models. Must be a sigverse-style data.frame (contain colums: channel, type, fraction, count). See \href{https://github.com/selkamand/sigshared?tab=readme-ov-file#sigverse-data-types}{sigshared readme} for details.
```

*Signature*

```         
signature: The profile of a mutational signature. Must be a data.frame (with columns: type, channel & fraction). See \href{https://github.com/selkamand/sigshared?tab=readme-ov-file#sigverse-data-types}{sigshared readme} for details.
```

*Signature collection*

```         
signatures:     A sigverse signature collection. A named list of sigstash signature data.frames. See \href{https://github.com/selkamand/sigshared?tab=readme-ov-file#sigverse-data-types}{sigshared readme} for details.
```

*Catalogue collection*

```         
catalogues: A sigverse-style catalogue collection. A named list of sigstash catalogue data.frames. See \href{https://github.com/selkamand/sigshared?tab=readme-ov-file#sigverse-data-types}{sigshared readme} for details.
```

*Cohort Signature Analysis Results*

```         
cohort: A sigverse-style data.frame describing the contributions of signatures in each sample (must contain columns: sample, signature contribution_absolute, contribution). See \href{https://github.com/selkamand/sigshared?tab=readme-ov-file#sigverse-data-types}{sigshared readme} for details.
```

*Signature Model Specification*

```         
model: A named numeric vector where names represent signatures and values represent their proportional contribution to the model. See \href{https://github.com/selkamand/sigshared?tab=readme-ov-file#sigverse-data-types}{sigshared readme} for details.
```

## 

## Example Data

For each data structure used by the sigverse we include a toy example.

```{r}
example_annotations()
example_bootstraps()

example_catalogue()
example_catalogue_collection()

example_signature()
example_signature_collection()

example_cohort_analysis()
```

We also include examples from real SBS mutational signature analysis of
the colo829 melanoma cell line

```{r paged.print=TRUE}
head(example_catalogue_colo829())
head(example_bootstraps_colo829())
```

## Other Utility Functions

sigverse dependencies are kept minimal. To mitigate the cost to
readability, sigshared includes some baseR implementations of common
data.frame manipulation functions and other utilities.

```{r}
# Setup data
mtcars <- head(mtcars, n = 5)

# Rename a dataframe
brename(mtcars, c(miles_per_gallon = "mpg"))

# Select a subset of columns
bselect(mtcars, c("mpg"))

# Evaluate code with a specific random seed
with_seed(seed = 123, { runif(1) })
```
