---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sigshared

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/sigshared)](https://CRAN.R-project.org/package=sigshared)

<!-- badges: end -->

**sigshared** contains general utilities used in many other R packages in the [sigverse](https://github.com/selkamand/sigverse)

Unless you're developing / maintaining a sigverse package, there is probably very little reason to ever download this package explicitly. We reccomend just installing the full sigverse as described [here](https://github.com/selkamand/sigverse)

## Installation

You can install the development version of sigshared from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("selkamand/sigshared")
```

## Sigverse Data Types

+---------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Data Structure                        | Requirements                                                                                                                                                                                    |
+=======================================+=================================================================================================================================================================================================+
| **Signature**                         | data.frames with 3 columns                                                                                                                                                                      |
|                                       |                                                                                                                                                                                                 |
|                                       | 1.  **type**                                                                                                                                                                                    |
|                                       | 2.  **channel**                                                                                                                                                                                 |
|                                       | 3.  **fraction**                                                                                                                                                                                |
+---------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Signature Collections**             | Lists of signature data.frames, where name of list entry is the name of the signature                                                                                                           |
+---------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Signature Annotations**             | Signature level annotations. data.frames with 4 required columns:                                                                                                                               |
|                                       |                                                                                                                                                                                                 |
|                                       | 1.  **signature**                                                                                                                                                                               |
|                                       | 2.  **aetiology**                                                                                                                                                                               |
|                                       | 3.  **class**                                                                                                                                                                                   |
|                                       | 4.  **subclass**                                                                                                                                                                                |
|                                       |                                                                                                                                                                                                 |
|                                       | class and subclass of aetiology do not need to conform to any specific ontology. However, we include the data-dictionary used by sigstash collections below (see [Signature Aetiology Classes]) |
+---------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Decompositions**                    | data.frames with 4 columns                                                                                                                                                                      |
|                                       |                                                                                                                                                                                                 |
|                                       | 1.  **channel**                                                                                                                                                                                 |
|                                       | 2.  **type**                                                                                                                                                                                    |
|                                       | 3.  **fraction**                                                                                                                                                                                |
|                                       | 4.  **count**                                                                                                                                                                                   |
+---------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Decomposition Collections**         | Lists of decomposition data.frames (1 per sample) where name represents a sample identifier.                                                                                                    |
+---------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Cohort Signature Analysis Results** | data.frame with 4 columns:                                                                                                                                                                      |
|                                       |                                                                                                                                                                                                 |
|                                       | 1.  **sample**                                                                                                                                                                                  |
|                                       |                                                                                                                                                                                                 |
|                                       | 2.  **signature**                                                                                                                                                                               |
|                                       |                                                                                                                                                                                                 |
|                                       | 3.  **contribution_absolute**                                                                                                                                                                   |
|                                       |                                                                                                                                                                                                 |
|                                       | 4.  **contribution (**percentage**)**                                                                                                                                                           |
|                                       |                                                                                                                                                                                                 |
|                                       | 5.  **bootstraps** (colon separated string)                                                                                                                                                     |
+---------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

### Assertions

You can assert an object belongs to each of data types

```{r, eval = FALSE}
library(sigshared)

# Generate Example Datatypes

# Signatures
signature = example_valid_signature()
signature_collection = example_valid_signature_collection()
signature_annotations = example_valid_signature_annotations()

# Decompositions
decomposition = example_valid_decomposition()
decomposition_collection = example_valid_decomposition_collection()

# Cohort Analysis Results
cohort_analysis = example_valid_cohort_analysis()


# Run Assertions
assert_signature(signature)
assert_signature_collection(signature_collection)
assert_decomposition(decomposition)
assert_decomposition_collection(decomposition_collection)
assert_cohort_analysis(cohort_analysis)
signature_annotations(decomposition_collection)
```

### Signature Aetiology Classes {data-link="Signature Aetiology Classes"}

| Class                         | Subclass                            |
|-------------------------------|-------------------------------------|
| artefact                      | 8-oxo-guanine                       |
| artefact                      | sequencing_artefact                 |
| artefact                      | germline_contamination              |
| artefact                      | oversegmentation                    |
| clock-like                    | clock-like                          |
| dysfunctional_dna_repair      | MMR                                 |
| dysfunctional_dna_repair      | HR                                  |
| dysfunctional_dna_repair      | NER                                 |
| dysfunctional_dna_repair      | BER                                 |
| dysfunctional_dna_repair      | NHEJ                                |
| dysfunctional_dna_replication | proofreading                        |
| dysfunctional_dna_replication | polymerase_mutations                |
| treatment_associated          | chemotherapy_platinum               |
| treatment_associated          | chemotherapy_thiopurine             |
| treatment_associated          | chemotherapy_pyrimidine_antagonists |
| treatment_associated          | chemotherapy_unknown                |
| treatment_associated          | chemotherapy_nitrogen_mustards      |
| treatment_associated          | triazenes                           |
| treatment_associated          | immunosuppression                   |
| environmental_mutagens        | tobacco                             |
| environmental_mutagens        | haloalkanes                         |
| environmental_mutagens        | UV                                  |
| plants_and_microbes           | aflatoxin                           |
| plants_and_microbes           | colibactin                          |
| plants_and_microbes           | aristolochic_acid                   |
| cytidine_deaminases           | cytidine_deaminases                 |
| immune                        | ROS                                 |
| dysfunctional_epigenetics     | topology                            |
| chromosomal                   | chromosomal_losses                  |
| chromosomal                   | chromosomal_instability             |
| chromosomal                   | chromothripsis                      |
| ploidy                        | diploid                             |
| ploidy                        | tetraploid                          |
| unknown                       | unknown                             |

### 
