---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sigshared

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/sigshared)](https://CRAN.R-project.org/package=sigshared)
<!-- badges: end -->

**sigshared** contains general utilities used in many other R packages in the [sigverse](https://github.com/selkamand/sigverse)

Unless you're developing / maintaining a sigverse package, there is probably very little reason to ever download this package explicitly. We reccomend just installing the full sigverse as described [here](https://github.com/selkamand/sigverse)


## Installation

You can install the development version of sigshared from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("selkamand/sigshared")
```

## Toolkits

### Assertions for key data structures

```{r, eval = FALSE}
assert_signature()
assert_signature_collection()
assert_decomposition()
assert_decomposition_collection()
assert_cohort_analysis()
```

